On the Dynamics of Anthropomorphism in Robotics

### 2013-07-01 dynamics-anthropomorphism.tex
WIP on the Familiarization phase

\paragraph{Disruptive behaviors}
\ref{sec:disruptive}

@@

>>>>

\emph{Familiarization}, in this context, means \emph{to get acquainted with}:
it is not to be confused with the \emph{familiarity thesis} that relates to the
projection of already known cognitive models onto the robot. We discuss this
aspect at section~\ref{sec:cognitivemodel}.

The \emph{familiarization phase} starts at the peak of the \emph{novelty
effect}: anthropomorphic effects are at their maximum, the user thinks (s)he is
\emph{potentially} facing an agent endowed with ``human-level intelligence'' to
take McCarthy's words.

This is a transient state that vanishes as soon as the human observe that the
behaviour of the robot is \emph{predictable} and possibly \emph{deceptive} (for
instance, it talks but does not understand when we talk to him; it has eyes but
does not recognize everyday objects I show to him; etc.).

\paragraph{Disruptive behaviors}

The model also represents \emph{disruptive behaviors} as local
increases of anthropomorphic effects in the familiarization phase (and at a
lesser extend, in the stabilization phase) (Figure~\ref{fig:dynamics}).

By \emph{disruptive behaviors}, we mean any behavior exhibited by the robot
that is unexpected by the user: for instance, a robot may usually follow always
the same route to go to two different places in a house, and suddenly change
the route. The actual underlying reason may span from a bug to the detection of
a new possible obstacle, but as long as this reason is not immediately
understood by the user, we count the behavior as \emph{disruptive}.

Section~\ref{sec:disruptive} propose a typology of disruptive behaviors and
discusses the possible consequences on the level of anthropomorphism, which may
both increase if the user perceives (rightfully or not) a form of
intentionality, or decrease if the unexpected behavior is perceived as a
failure.

@@

Besides, we propose to call \emph{expectation mismatch} the delta between the
initial capital of anthropomorphism for a given tuple (user, robot, context)
and the anthropomorphism peak of novelty effect (Figure~\ref{fig:dynamics}).
\fixme{Need to rediscuss what this delta actually represent...}
